Image projections along an axis

### 2006-03-03 article/Article.tex
continue article
darcs-hash:20060303132903-2fc9d-f8541dc931fe1ae82df0cddc2898fbd8be6394e5.gz

\section{Introduction}
ITK already provide a class to project an image along an axis:
\doxygen{AccumulateImageFilter}. However, this class is limited to sum
and mean projections.

@@

The base class, \verb$ProjectionImageFilter$, is mainly a copy of
\verb$AccumulateImageFilter$, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
  \item thread support
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (\verb$MaximumProjectionImageFilter$)
  \item minimum projection (\verb$MinimumProjectionImageFilter$)
  \item mean projection (\verb$MeanProjectionImageFilter$)
  \item median projection (\verb$MedianProjectionImageFilter$)
  \item sum projection (\verb$SumProjectionImageFilter$)
  \item sigma projection (\verb$SigmaProjectionImageFilter$)
  \item binary projection (\verb$BinaryProjectionImageFilter$)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum2D0}
\caption{MaximumProjectionImageFilter, projection along axis x.\label{maximum0}}
\end{figure}

>>>>

\section{Introduction}
Image projection is a very common task in image analysis to reduce the dimension
of an image. ITK already provide a class to project an image along an axis:
\doxygen{AccumulateImageFilter}. However, this class is not very efficient,
hard to extend, and not able to produce an image with a smaller dimension than
the input image.

Like \verb$AccumulateImageFilter$, \verb$ProjectionImageFilter$ can produce an
ouput image with the same dimension than the input image. In that case, all
the informations associated to the input image are copied to the output one,
and excepted:
\begin{itemize}
  \item the size of the image which is set to 1 on the projection dimension,
  \item the spacing of the image, which is modified on the projection dimension
to keep the physical size of the image on the output image,
  \item the index of the image wich is set to 0 on the projection dimension.
\end{itemize}
Basically, \verb$ProjectionImageFilter$ produce the same volume but with a
only 1 pixel on the projection dimension.

Unlike the \verb$AccumulateImageFilter$, \verb$ProjectionImageFilter$ can 
also produce an ouput image with a dimension lower than the input image.
While it is logical to produce a 2D image with a projection of a 3D image,
it also bring some more complex transforms, because the dimensions of the
input image can't be directly copied to the input image - one dimension is
removed. \verb$ProjectionImageFilter$ move the removed dimension of the
input image to the projected dimension in the output image. With a 3D
input image, the output will be:
\begin{itemize}
  \item projection along the axe x:
	\begin{itemize}
	\item x is removed
	\item y stay y in the output image
	\item z become x in the output image
	\end{itemize}
  \item projection along the axe y:
	\begin{itemize}
	\item x stay x in the output image
	\item y is removed
	\item z become y in the output image
	\end{itemize}
  \item projection along the axe z:
	\begin{itemize}
	\item x stay x in the output image
	\item y stay y in the output image
	\item z is removed.
	\end{itemize}
       No dimension is moved in that case.
\end{itemize}

Also, during the projection, there is an orientation problem which doesn't
exist if the number of dimensions is kept.

\small \begin{verbatim}
        ________
       /       /|            /|
      /       / |           / |
 1   /_______/  |  2       /  |
 --> |       |  |<--       |  |
     |       | 0/          | 0/
     |       | /           | /
     |_______|/            |/
\end{verbatim} \normalsize

On the volume above, the image is projected on the horizontal dimension.
If the number of dimension is not modified, a thin volume with a size of
1 on projection dimension, and there is no problem.

If the number of dimensions is reduced, the direction become important,
and the result is not the same.

\small \begin{verbatim}
      _______
     |0      |
 1   |       |
     |_______|
      _______
     |      0|
 2   |       |
     |_______|
\end{verbatim} \normalsize

However, we choose to move the dimensions and the indexes on those
dimensions, and so to not provide the ability to choose the direction.
The image is always projected with direction 1.
The user can still modify the output image with ITK filters like
\doxygen{FlipImageFilter}.

@@

The base class, \verb$ProjectionImageFilter$, is a re-write of
\verb$AccumulateImageFilter$, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
  \item thread support
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (\verb$MaximumProjectionImageFilter$)
  \item minimum projection (\verb$MinimumProjectionImageFilter$)
  \item mean projection (\verb$MeanProjectionImageFilter$)
  \item median projection (\verb$MedianProjectionImageFilter$)
  \item sum projection (\verb$SumProjectionImageFilter$)
  \item sigma projection (\verb$SigmaProjectionImageFilter$)
  \item binary projection (\verb$BinaryProjectionImageFilter$)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum2D0}
\caption{MaximumProjectionImageFilter, projection along axis x. Note that the orientation
is the correct one: the axis z (34 pixels) have been moved to the axis x, and the axis y
(371 pixels) is not modified. Also, the image shown here does not respect the spacing
of the image, but ProjectionImageFilter does.\label{maximum0}}
\end{figure}
